---
title: "Gautam Kapila Case Study 1"
author: "G.Kapila"
date: "7/5/2019"
output:
  html_document:
  keep_md: yes
---

# United States Brewery Market and Beer Composition Statistics

## Introduction 

The work summarizes following aspects of US brewery market and beer composition per state

* Number of breweries present per state
* Acoholic content of craft beers per state
* Bitterness of the beers per state
* State with beer having maximum alchoholic content and bitterness
* Relationship between beer alcoholic content and beer bitterness

It is envisaged that this work could be extended in future, to come up with proposals on where to start a new brewery venture and target beer composition


### Session and Library Information

```{r documentation, echo=TRUE}

sessionInfo()

library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)

```
### Data Preparation
```{r readData, echo=TRUE}
beers <- read.csv(file = "CaseStudy1_2_2_2\\beers.csv",sep = ",")
breweries <- read.csv(file = "CaseStudy1_2_2_2\\breweries.csv",sep = ",")
bd    <- merge(beers, breweries, by.x = "Brewery_id", by.y = "Brew_ID")
bd$State <- gsub('\\s+', '', bd$State)
colnames(bd) <- c("BreweryID","BeerName","BeerID","ABV","IBU","Style","Oz","BreweryName","City","State")
write.csv(bd,file = "bd.csv",row.names = FALSE)
```

### Breweries present in each state
```{r breweryDistribution, echo=TRUE, message=FALSE, warning=FALSE}
BDist <- bd %>% dplyr::count(State,sort = TRUE, name = "Count")
ggplot(data = BDist, aes(x=reorder(State,Count),y=Count,fill = Count)) +
    geom_bar(stat = "identity",width = 0.7) +
    coord_flip() +
    geom_text(aes(label=Count), hjust = -0.5, vjust = 0.1, size=2)+
    ggtitle('Brewery Count by State') +
    theme(plot.title = element_text(hjust = 0.5)) +
    xlab("State")+
    theme(axis.text = element_text(size = 8)) +
    scale_fill_gradient(low = "green",high = "red")

  

```

### Printing first 6 and last 6 data from merged file

```{r mergedData, echo=TRUE, message=FALSE, warning=FALSE}

rbind(head(bd,6),tail(bd,6))

```

### Printing NA's count per column

```{r naCount, echo=TRUE, message=FALSE, warning=FALSE}
sapply(bd, function(x) sum(length(which(is.na(x)))))
```

### Median Alcohol Content and IBU for each state

```{r medianAlcohoAndIBU, echo=TRUE, message=FALSE, warning=FALSE}

mABVbyState <- bd %>% drop_na(ABV) %>% group_by(State) %>% summarize(mABV = median(ABV))
ggplot(data = mABVbyState, aes(x=reorder(State,mABV),y=mABV,fill = mABV)) +
    geom_bar(stat = "identity",width = 0.7) +
    coord_flip() +
    geom_text(aes(label=mABV), hjust = -0.5, vjust = 0.1, size=2)+
    ggtitle('Median ABV by State') +
    theme(plot.title = element_text(hjust = 0.5)) +
    xlab("State")+
    ylab("Median ABV")+
    theme(axis.text = element_text(size = 8)) +
    scale_fill_gradient(low = "green",high = "red") +
    ylim(0,0.07)


mIBUbyState <- bd %>% drop_na(IBU) %>% group_by(State) %>% summarize(mIBU = median(IBU))
ggplot(data = mIBUbyState, aes(x=reorder(State,mIBU),y=mIBU,fill = mIBU)) +
    geom_bar(stat = "identity",width = 0.7) +
    coord_flip() +
    geom_text(aes(label=mIBU), hjust = -0.5, vjust = 0.1, size=2)+
    ggtitle('Median IBU by State') +
    theme(plot.title = element_text(hjust = 0.5)) +
    xlab("State")+
    ylab("Median IBU")+
    theme(axis.text = element_text(size = 8)) +
    scale_fill_gradient(low = "green",high = "red")

```

### State with Max Alcohol Content and Max IBU 
#### State with Max Alcoholic Beer

```{r maxAlcohol, echo=TRUE, message=FALSE, warning=FALSE}

maxABVbyState <- bd %>% drop_na(ABV) %>% group_by(State) %>% summarize(mABV = max(ABV))
maxABVState <- as.character(maxABVbyState[order(-maxABVbyState$mABV),][1,1])
maxABVState
```

#### State with most bitter bear (max IBU)
```{r maxIBU, echo=TRUE, message=FALSE, warning=FALSE}

maxIBUbyState <- bd %>% drop_na(IBU) %>% group_by(State) %>% summarize(mIBU = max(IBU))
maxIBUState <- as.character(maxIBUbyState[order(-maxIBUbyState$mIBU),][1,1])
maxIBUState

```

### Summary Statistics for ABV variable
```{r summaryABV, echo=TRUE, message=FALSE, warning=FALSE}

summary(beers$ABV)

```

### Relationship between Beer Bitterness and Alcoholic Content
```{r scatterPlot, echo=TRUE, message=FALSE, warning=FALSE}

ggplot(bd, aes(x=ABV, y=IBU, color=State)) +
  geom_point() +
  xlab("Alcoholic Content of Beer, ABV")+
  ylab("Bitterness of Beer, IBU") +
  theme(axis.text = element_text(size = 8)) +
  ggtitle('Scatter plot between bitterness and alcoholic content') +
  theme(plot.title = element_text(hjust = 0.5))  

```

#### Following aspects of the relationship are evident
##### 1. In general, higher bitterness tracks with higher alcholic content.
##### 2. It appears that one can have a range of bitterness for a given alcoholic content, and vice versa. 
#####    There is more to getting the right bitterness (IBU) for a beer than just it's alcoholic content (ABV)

## Conclusions

* Colorado by far has the largest number of breweries. 
    + CA, MI, IN, TX are other states with high density of breweries.
    + _Targeting these states for additional opening of brewery should be considered._
* Co ranks 5th in list of states with highest alcholol content producing states.
    + However, it produces beer with highest alcholic content.
    + _Perhaps there is additional apetite in market to absorb higher alcholic content beer._
* Clearly bitterness and alcoholic content track each other.
    + However, beers with different amount of bitterness can be produced at any given alcoholic content.
    + _One can be flexible in tailoring beer composition to suite whats popular in local market_
    
## GitHub Reference

https://github.com/gkapila07/msds_homeworks/tree/master/dds 